\section{Introduction-Xu}

A student asks a tutoring system, ``Why is my answer wrong?'' At first glance, the system may need to retrieve relevant course content or rubric guidance. But an effective tutor must do more than surface the right information. It must decide whether the student needs a direct explanation, a scaffolded hint, a rubric-based correction, or even a deliberate withholding of the full answer in order to support learning. What begins as an information request quickly becomes an instructional decision-making problem.

Retrieval-augmented generation has become the standard way to ground language model outputs in external evidence~\cite{lewis2020rag}, and that default works well for fact-based question answering. In tutoring, however, grounded retrieval is only one part of the interaction. A tutoring system may also need to infer learner state, decide how much help to provide, align with pedagogical goals, and determine whether instructional strategy should take precedence over direct answer delivery. The challenge is therefore not whether retrieval is useful, but when retrieval is sufficient and when additional tutoring control is required.

Recent education-centered benchmarks make this tension more visible. Datasets such as EduBench and TutorEval assess not only correctness but also pedagogy, adaptivity, reasoning, and grounded tutoring behavior~\cite{xu2025edubench,chevalier2024sciencetutors,lmsciencetutorrepo}. Related benchmarks and evaluation studies further emphasize pedagogical quality, adaptivity, and safety rather than answer accuracy alone~\cite{macina2025mathtutorbench,srinivasa2025tutorbench,maurya2025aitutoreval,borchers2025adaptivity,hazra2026safetutors,shi2026kmpbench}. In these settings, correctness remains essential, but it is no longer sufficient. These shifts in evaluation criteria make the architectural question more concrete: if a task is constrained mainly by evidence access, then classical RAG may remain the most appropriate top-level design; if the main bottleneck lies instead in diagnosis, planning, routing, rubric following, or targeted pedagogical intervention, then more explicit orchestration may matter more than retrieval alone.

This paper studies a simple question---when is retrieval needed, and when is orchestration justified?---through an adaptive multi-agent tutoring framework with regime routing and conditional retrieval. Here, orchestration refers to the additional control needed to organize tutoring behavior, including diagnosis, planning, routing, critique, and selective intervention beyond retrieval alone. Instead of assuming that one architecture should govern all educational tasks, the framework routes tasks among different execution modes, including classical RAG, a hybrid agentic path, and coordination-heavy non-RAG tutoring. Retrieval is used when evidence is needed, but it is not treated as the default controller in every situation. The central idea is straightforward: the preferred architecture may vary across educational regimes because different tutoring tasks are shaped by different combinations of evidence requirements and pedagogical coordination demands. This motivation is also consistent with recent multi-agent work suggesting that structured coordination can outperform a single undifferentiated prompting strategy once tasks require decomposition, planning, or staged reasoning~\cite{nguyen2025marag,radeva2025coordination}.

To ground this framing empirically, we report preliminary repository-grounded experiments on aligned 30-example snapshots from EduBench and TutorEval. These experiments are not intended as definitive benchmark results, but as early tests of the paper's central question: whether architecture preference remains stable across educational settings or instead shifts with task regime. By comparing classical RAG, hybrid agentic, and coordination-heavy tutoring routes under aligned snapshot conditions, the experiments provide an initial way to observe when retrieval-centered designs appear sufficient and when more explicit orchestration becomes justified. Our goal is therefore comparative rather than doctrinal. We do not claim that multi-agent systems universally replace retrieval. Instead, we use these preliminary experiments to examine whether preferred architectures change with differences in pedagogical and evidence demands, and whether retrieval is most effective as a conditional capability within a broader coordination strategy.

The broader significance of this question is both practical and methodological. Within education, it suggests that tutoring systems should not be designed as if better grounding alone guarantees better instructional support. They must also account for diagnosis, scaffolding, rubric alignment, selective retrieval, and the cost of additional coordination. More broadly, although this study is grounded in educational tutoring, the underlying design logic may also help inform architecture selection in other settings where grounding, orchestration, and efficiency must be balanced together. In that sense, the contribution of this paper is not to declare a universal winner, but to show that conditional retrieval and adaptive orchestration can provide a more informative design perspective than retrieval-first assumptions alone.

\section{Introduction}

Large language model systems for education are frequently characterized as retrieval-augmented generation (RAG) pipelines: an inquiry submitted by a user is matched to an external knowledge source, relevant passages are retrieved, and the model produces a substantiated response. This methodology proves effective, particularly when the task is predominantly factual, citation-sensitive, or directly answerable from curated course materials. Nevertheless, numerous educational tasks extend beyond mere information retrieval. Tutoring, rubric-based feedback, misconception diagnosis, and study planning necessitate multi-step pedagogical coordination, adaptation to the learner's state, and iterative refinement, rather than simple retrieval.

This paper inquires into a more specific issue than the general question of whether multi-agent systems should supplant retrieval-based systems in educational contexts. Instead, it examines the operational conditions under which the classical Retrieval-Augmented Generation (RAG) approach suffices, and when systems requiring more extensive orchestration become justified. Within this framework, retrieval remains significant for tasks grounded in evidence; however, it may not always serve as the most informative top-level abstraction for tasks whose primary difficulty resides in pedagogical coordination. Accordingly, the objective is not to advocate for the universal superiority of multi-agent systems, but rather to delineate the boundary conditions under which retrieval-centric, orchestration-centric, or hybrid configurations are most suitable..

To support this objective, the paper categorizes educational AI tasks along two dimensions: the necessity for knowledge grounding and the complexity of pedagogical coordination. This framework enables the comparison of traditional RAG, agentic RAG, and non-RAG multi-agent systems not as mere interchangeable labels, but as distinct solutions tailored to varying task requirements. A straightforward retrieval process may better serve tasks characterized by minimal coordination and high knowledge grounding. In contrast, educational workflows requiring extensive coordination could warrant more explicit planning, critical assessment, or role specialization.

The paper's contributions are threefold. First, it proposes a task taxonomy that separates retrieval-heavy educational tasks from coordination-heavy ones. Second, it outlines a framework for comparing classical RAG, agentic RAG, and non-RAG multi-agent systems under shared budgets, source corpora, and evaluation protocols. Third, it identifies a benchmark and dataset strategy for evaluating architectural tradeoffs in educational AI without overstating downstream learner outcomes. The goal is not to argue that more agents are always better, but to clarify when orchestration complexity is justified and when classical retrieval-centric pipelines remain the right design choice.

\subsection{Research Questions}

The paper is organized around the following research questions.(Xu Rewrite to 2 RQs)

\begin{enumerate}[label=RQ\arabic*:,leftmargin=*]
    \item Under what educational task conditions is classical RAG sufficient, and when do agentic or multi-agent systems become justified?
    \item What evaluation framework can fairly compare classical RAG, agentic RAG, and non-RAG multi-agent systems under shared tasks, budgets, and source corpora?
\end{enumerate}

\subsection{Hypotheses}

To keep the framing falsifiable, the paper adopts balanced hypotheses rather than a universal superiority claim.

\begin{enumerate}[label=H\arabic*:,leftmargin=*]
    \item On low-coordination, knowledge-grounded educational tasks, classical RAG will perform comparably to or better than more complex multi-agent systems at lower cost.
    \item On coordination-heavy educational tasks such as adaptive tutoring, rubric-based critique, and multi-step study planning, agentic or multi-agent systems may outperform classical RAG on pedagogical quality and adaptivity.
    \item Non-RAG multi-agent systems will underperform retrieval-enabled systems on fact-intensive or citation-sensitive tasks, showing that orchestration alone does not remove the need for grounding.
    \item A plausible hypothesis for mixed-regime educational tasks is that hybrid systems will prove strongest: multi-agent orchestration with retrieval available as a conditional tool rather than as the sole pipeline pattern.
\end{enumerate}
