\documentclass[11pt]{article}

\usepackage[margin=1in]{geometry}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{lmodern}
\usepackage{microtype}
\usepackage{hyperref}
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage{multirow}
\usepackage{array}
\usepackage{enumitem}
\usepackage{xcolor}

\title{Adaptive Tutoring Beyond Classical RAG:\\Task Taxonomy, Evaluation Design, and Repository-Grounded Multi-Agent Evidence}
\author{Draft in Progress}
\date{\today}

\begin{document}

\maketitle

\begin{abstract}
Classical retrieval-augmented generation (RAG) is a strong default for evidence-grounded educational question answering, but many tutoring tasks depend on diagnosis, scaffold selection, rubric alignment, and interaction planning in addition to document access. This paper asks when those coordination demands justify treating retrieval as a conditional tool inside a broader multi-agent tutoring policy rather than as the top-level system design. We make three contributions. First, we present a task taxonomy that separates educational workloads by knowledge-grounding need and pedagogical coordination complexity. Second, we outline a comparison framework for classical RAG, agentic RAG, and non-RAG multi-agent systems that keeps groundedness, pedagogical quality, adaptivity, robustness, cost, and latency distinct under shared budgets and shared task conditions. Third, we incorporate repository-grounded evidence from article-scale snapshots on EduBench and TutorEval to show how the preferred architecture changes with benchmark demands: coordination-heavy settings favor non-RAG or hybrid orchestration, while long-context evidence-grounded tutoring still benefits from conditional retrieval. The resulting claim is disciplined rather than universal. In education, multi-agent systems are best understood not as a blanket replacement for retrieval, but as a stronger orchestration option when pedagogical coordination becomes the main systems problem.
\end{abstract}

\input{sections/introduction}
\input{sections/related_work}
\input{sections/framework}
\input{sections/evaluation}
\input{sections/results}
\input{sections/analysis}
\input{sections/discussion}
\input{sections/conclusion}
\input{sections/appendix}

\bibliographystyle{plain}
\bibliography{references/references}

\end{document}
